---
title: "Eating Out"
author: "Kevin Brandon Salazar"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this analysis we will figure out if there is a difference in the frequency of eating out based on ethnicity

## There were two years in which the General Social Survey asked about the frequency of eating out.
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
```
```{r, eval=TRUE}
GSS <-readstata13::read.dta13(file="GSS1996.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

# Method
To make sense of the data from 1996 provided by gss.NORC.org about which ethnicity eats out most frequently needs to be determined using a categorical numerical method with race being categorical and frequency being numerical.

# Results

## Descriptive Results

### Graphical Results

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata,type="percent")

```

The bar chart suggests that there are differences in the proportions of white and other ethnic groups who eat out displaying that white race eats out more frequent than others by a small margin. 

### Numerical Description

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```
The results in this numerical summary shows that whites eat out more frequently compared to other ethnicities.

### Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisherXY(mydata$race,mydata$eatout,printAll=T)
```
The Chi-squared test shows the P-value is 0.0008179, and the Fisher Exact test for count data has a P-value of 0.001208 so we see that the number is so small we can reject the null hypothesis.  

# Discussion/Conclusions

Based on the Fisher Exact test we only have evidence that there's a difference in eating out based on white and black race for the P-value is so small it shows that there exists a difference (P=0.008179) so I reject the null hypothesis. While we have no evidence about difference in the frequency of eating out for whites and other because the p value (0.35) is so large we fail reject the null hypothesis. While we also have no evidence about difference in the frequency of eating out for blacks and other because the p value (0.28) is so large we also fail reject the null hypothesis.
